---
title: "Feature_extraction_pipeline_using_EBImage"
author: "Huayu"
date: "2019 M08 14"
output:
  pdf_document: default
  html_document: default
---

# Env setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# Librarys
required_libs <- c("tiff", "ggplot2", "png", "EBImage", "raster")

new_libs <- required_libs[!(required_libs %in% installed.packages()[,"Package"])]

if (length(new_libs) > 0) install.packages(new_libs, dependencies = T, quiet = T)

new_libs <- required_libs[!(required_libs %in% installed.packages()[,"Package"])]

if (length(new_libs) > 0)
{
  if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager", quiet = T)
  
  BiocManager::install(new_libs)
}


for (i in 1:length(required_libs))
  library(required_libs[i], character.only = T)

rm(list = c("i", "new_libs", "required_libs"))

sessionInfo()


# Function definiations
img.Import <- function(index, fileList) # Import image file and assemble the list
{
  imgRGB <- readTIFF(source = fileList$RGB[index])
  imgYGB <- readTIFF(source = fileList$YGB[index])
  
  return(list(
    R = imgRGB[,,1],
    G = imgRGB[,,2],
    B = imgRGB[,,3],
    Y = imgYGB[,,1]
  ))
}

img.Import_512 <- function(index, fileList) # Import image file and assemble the list
{
  return(
  
  list(
    "R" = readPNG(source = fileList$R[index]),
    "G" = readPNG(source = fileList$G[index]),
    "B" = readPNG(source = fileList$B[index]),
    "Y" = readPNG(source = fileList$Y[index])
  )
  )
  
}

binChannel <- function(channel) # Binarize channels using the background color plus 5/255
{
  step <- c()
  
  for (i in 1:30)
  {
    step[i] <- sum(channel < i/255 & channel >= (i-1)/255)
  }
  
   
  return((which.max(step) + 5)/255)
}


Nb8 <- as.matrix(expand.grid(x = -1:1, y = -1:1))


nbEight <- function(x, minEdge = 0, maxEdge = 513) { # Get neightbours of positive area
  n <- nrow(x)
  series <- rep(1:n, each = 9)
  nbs <- x[series,]
  nbs[,1] <- nbs[,1] + Nb8[,1]
  nbs[,2] <- nbs[,2] + Nb8[,2]
  
  nbs <- unique(nbs)
  nbs <- nbs[(nbs[,1] > minEdge) & (nbs[,1] < maxEdge) &  (nbs[,2] > minEdge) & (nbs[,2] < maxEdge) ,]
  
  
  nbs
}

rimOfNuclei <- function(nucMask) # Get rim of nuclues
{
  rims <- do.call(rbind, ocontour(nucMask))
  n <- nrow(nucMask)
  
  if (n == 0) 
  
  rims <- nbEight(rims)
  rimsLin <- (rims[,2] - 1) * n + rims[,1]
  
  rimMask <- rep(0, n^2)
  
  rimMask[rimsLin] <- 1
  rimMask <- matrix(rimMask, nrow = n)
  
  rimMask

}

huayu.cutoff <- function(img)
{
  if (sum(img[img > 0]) < 50)
    return(0.01)
  
  img_hist <- hist(img[img > 0], breaks = 50, plot = F)$density
  n <- length(img_hist)
  
  img_hist_d1 <- img_hist[2:n] - img_hist[1:(n-1)]
  img_hist_d2 <- img_hist_d1[2:(n-1)] - img_hist_d1[1:(n-2)]
  
  img_hist_d2_fluc <- abs(img_hist_d2) > 0.05
  
  for (i in (n-3):10)
    if (sum(img_hist_d2_fluc[(i-1):(i+1)]) == 3)
      if (img_hist_d2[i-1] > 0 & img_hist_d2[i] < 0 & img_hist_d2 [i + 1]> 0)
        return(i * 0.02)
  
  for (i in (n-3):1)
    if (img_hist_d2_fluc[i]) return(i * 0.02)
  
  return(quantile(img, prob = 0.95))
}

coloc <- function(bin1, bin2)
{
  
  
  if (sum(bin1) == 0 | sum(bin2) == 0 ) return(0)
  
  both <- sum(bin1 & bin2)
  
  ratio1 <- both/sum(bin1)
  ratio2 <- both/sum(bin2)
  
  colocScore <- 2*ratio1*ratio2/(ratio1 + ratio2)
  
  return(colocScore)
}


# File list

img_list <- list(
  RGB = list.files(path = "./example_img/img/RGB/", full.names = T),
  YGB = list.files(path = "./example_img/img/YGB/", full.names = T)
)

img_512_list = list(
  R = list.files(path = "./example_img/img_512/", pattern = "red", full.names = T),
  G = list.files(path = "./example_img/img_512/", pattern = "green", full.names = T),
  B = list.files(path = "./example_img/img_512/", pattern = "blue", full.names = T),
  Y = list.files(path = "./example_img/img_512/", pattern = "yellow", full.names = T)
)


```


# Image import

Import both original resolution image and image with reduced sise (512 by 512)
Immunofluorescent (IF) images are list of 4 channels: Red, Green, Blue and Yellow
Green channel is the information channel and the other channels are reference channel: Red (F-actin), Blue (DNA), Yellow (Mitochondria)
Resolution of orignal image is recorded

```{r}
img <- img.Import(index = 1, fileList = img_list)
img_512 <- img.Import_512(index = 1, fileList = img_512_list)
    
wh <- dim(img$R)[1]

print("structure of img is:")
str(img)

print("sturcture of img_512 is:")
str(img_512)
```

Preview of the channels

```{r}
display(img_512$R)
display(img_512$G)
display(img_512$B)
display(img_512$Y)
```


# Object mask creation

This section aims to create the masks for areas and objects in the IF image
including:
 * Nuclei
 * Cytoplasm
 * Extracellular area

 
## Image binarization

Binarize image on cutoff 5/255 higher than background color. Binarized images are used for finilizing the masks. For computation efficiency, low resolution image is used. 
 * DNA reference makes nucleus region
 * F-actin and mitochondria conbined for cytoplasm region

```{r}
img_cutoffs <- sapply(img_512, binChannel)

img_bin_512 <- list(
      "R" = img_512$R > img_cutoffs["R"],
      "G" = img_512$G > img_cutoffs["G"],
      "B" = img_512$B > img_cutoffs["B"],
      "Y" = img_512$Y > img_cutoffs["Y"]
    )

display(img_bin_512$B)
```

## Fillhull, watershed and remove noices

Fill the holes in large areas

```{r}
img_masks_512 <- list(
      "B" = fillHull(img_bin_512$B),
      "RY" = img_bin_512$R | img_bin_512$Y
    )

display(img_masks_512$B)
display(img_masks_512$RY)
```

Watershedding to define objects
Remove extra small objects, which are noise dots
Watershedding returns number matrix same size as the image. Indexes are given to objects, and number 0 is give to background.
Objects on the edge of the image are removed, becasue they are outlayers for many calculations.

```{r}
img_masks_512$B <- watershed(x = distmap(img_masks_512$B), tolerance = 1, ext = 3)
img_table_512 <- list(
      "B" = table(img_masks_512$B)
    )
img_masks_512$B <- rmObjects(x = img_masks_512$B, index = names(img_table_512$B)[img_table_512$B < 100])
img_table_512$B <- table(img_masks_512$B)

img_table_512$Bminus <- table(img_masks_512$B[2:511, 2:511])
img_masks_512$Bminus <- rmObjects(x = img_masks_512$B, index = names(img_table_512$B)[img_table_512$B > img_table_512$Bminus])

display(colorLabels(img_masks_512$B))
display(colorLabels(img_masks_512$Bminus))
    
```




```{r}
img_masks_512$RYinv <- bwlabel(1 - img_masks_512$RY)
img_table_512$RYinv <- table(img_masks_512$RYinv)
img_masks_512$RYinv <- rmObjects(x = img_masks_512$RYinv, index = names(img_table_512$RYinv)[img_table_512$RYinv > 20])
img_masks_512$RY[img_masks_512$RYinv > 0] <- 1
img_masks_512$RYinv <- NULL
    
img_masks_512$RY <- bwlabel(img_masks_512$RY)
img_table_512$RY <- table(img_masks_512$RY)
img_masks_512$RY <- rmObjects(x = img_masks_512$RY, index = names(img_table_512$RY)[img_table_512$RY < 100])
img_masks_512$RY[img_masks_512$B > 0] <- 0 

display(colorLabels(img_masks_512$RY))
```

Create mask for rim of nuclei, great mask for identifying nucleomembrane

```{r}
img_masks_512$rimB <- rimOfNuclei(nucMask = img_masks_512$Bminus)
display(x = img_masks_512$rimB)
```

## Upscale of the masks to full resolution
```{r}
img_masks <- lapply(img_masks_512, function(x, fact) as.matrix(disaggregate(x = raster(x), fact = fact)), fact = wh/512)
```


# Masks for signals in green channel

## Segmentation of green channel signals

Disect the green signal according to the masks (subcellular areas)

```{r}
img_Green <- list(
      "Nucleus" = img$G,
      "Cytosol" = img$G,
      "Rest" = img$G
    )
    
    img_Green$Nucleus[img_masks$B == 0] <- 0
    img_Green$Cytosol[img_masks$RY == 0] <- 0
    img_Green$Rest[img_masks$B > 0| img_masks$RY > 0 ] <- 0
    
    display(img_Green$Nucleus)
    display(img_Green$Cytosol)
    display(img_Green$Rest)
    
```


## Object masks for Green signals
Emprical way to deside cutoffs for green signal
Method 1: Otsu method which is more loose
Method 2: Right most fluctuation of 2nd derivative/Right most non-zero 2nd derivative, which is more stringent
For Nuclues and Cytosol use both methods
For Rest use Otsu only
Libraries with too many labels is a huge computation burden, sampling will be done

The otsu cutoff

```{r}
img_cutoffs <- c(img_cutoffs, 
                     G.Otsu = sapply(img_Green, otsu),
                     G.Huayu = sapply(img_Green[1:2], huayu.cutoff),
                     G.Otsu = otsu(img_512$G), 
                     G.Huayu = huayu.cutoff(img_512$G),
                     "R.Otsu" = otsu(img_512$R),
                     "B.Otsu" = otsu(img_512$B),
                     "Y.Otsu" = otsu(img_512$Y)
    )
    
    
img_masks_G.Otsu <- list(
      "Nucleus" = bwlabel(x = img_Green$Nucleus > img_cutoffs["G.Otsu.Nucleus"]),
      "Cytosol" = bwlabel(x = img_Green$Cytosol > img_cutoffs["G.Otsu.Cytosol"]),
      "Rest" = bwlabel(x = img_Green$Rest > img_cutoffs["G.Otsu.Rest"])
      
    )
    
img_table_G.Otsu <- lapply(X = img_masks_G.Otsu, FUN = table)
    
img_masks_G.Otsu$Nucleus <- rmObjects(
  x = img_masks_G.Otsu$Nucleus, 
  index = names(img_table_G.Otsu$Nucleus)[img_table_G.Otsu$Nucleus < 10])

if (sum(img_table_G.Otsu$Nucleus >= 10) > 200)
  img_masks_G.Otsu$Nucleus <- rmObjects(
    x = img_masks_G.Otsu$Nucleus, 
    index = names(img_table_G.Otsu$Nucleus)[sample(1:sum(img_table_G.Otsu$Nucleus >= 10), sum(img_table_G.Otsu$Nucleus >= 10) - 200)])
    
    
img_masks_G.Otsu$Cytosol <- rmObjects(
  x = img_masks_G.Otsu$Cytosol, 
  index = names(img_table_G.Otsu$Cytosol)[img_table_G.Otsu$Cytosol < 10])
if (sum(img_table_G.Otsu$Cytosol >=10) > 200)
      img_masks_G.Otsu$Cytosol <- rmObjects(
        x = img_masks_G.Otsu$Cytosol, 
        index = names(img_table_G.Otsu$Cytosol)[sample(1:sum(img_table_G.Otsu$Cytosol >= 10), sum(img_table_G.Otsu$Cytosol >= 10) - 200)])
    
img_masks_G.Otsu$Rest <- rmObjects(
      x = img_masks_G.Otsu$Rest, 
      index = names(img_table_G.Otsu$Rest)[img_table_G.Otsu$Rest < 10])
if (sum(img_table_G.Otsu$Rest >=10) > 200)
      img_masks_G.Otsu$Rest <- rmObjects(
        x = img_masks_G.Otsu$Rest, 
        index = names(img_table_G.Otsu$Rest)[sample(1:sum(img_table_G.Otsu$Rest >= 10), sum(img_table_G.Otsu$Rest >= 10) - 200)])

display(colorLabels(img_masks_G.Otsu$Nucleus))
display(colorLabels(img_masks_G.Otsu$Cytosol))


```

My customized cutoff
This more strigent cutoff captures objects with very strong signals, such as substructures in nuclei. 

```{r}
img_masks_G.Huayu <- list(
      "Nucleus" = bwlabel(x = img_Green$Nucleus > img_cutoffs["G.Huayu.Nucleus"]),
      "Cytosol" = bwlabel(x = img_Green$Cytosol > img_cutoffs["G.Huayu.Cytosol"])
    )
    
img_table_G.Huayu <- lapply(X = img_masks_G.Huayu, FUN = table)
    
    
if (sum(img_table_G.Huayu$Nucleus >= 10) > 100)
    img_masks_G.Huayu$Nucleus <- rmObjects(
      x = img_masks_G.Huayu$Nucleus, 
      index = names(img_table_G.Huayu$Nucleus)[order(img_table_G.Huayu$Nucleus, decreasing = T)[51:length(img_table_G.Huayu$Nucleus)]])
    
    
if (sum(img_table_G.Huayu$Cytosol >=10) > 100)
    img_masks_G.Huayu$Cytosol <- rmObjects(
      x = img_masks_G.Huayu$Cytosol, 
      index = names(img_table_G.Huayu$Cytosol)[order(img_table_G.Huayu$Cytosol, decreasing = T)[51:length(img_table_G.Huayu$Cytosol)]])
    
    
display(colorLabels(img_masks_G.Huayu$Nucleus))
display(colorLabels(img_masks_G.Huayu$Cytosol))
    
```


# Compute object features

Features are computed using
 * colocolization of green signal to reference channels
 * Object properties using masks


```{r}
manualFeatures <- c(
      "GcolocB" = coloc(bin1 = img$G > img_cutoffs["G.Otsu"], bin2 = img$B > img_cutoffs["B.Otsu"]),
      "GcolocR" = coloc(bin1 = img$G > img_cutoffs["G.Otsu"], bin2 = img$R > img_cutoffs["R.Otsu"]),
      "GcolocY" = coloc(bin1 = img$G > img_cutoffs["G.Otsu"], bin2 = img$Y > img_cutoffs["Y.Otsu"]),
      "GcolocRimB" = coloc(bin1 = img$G > img_cutoffs["G.Otsu"], bin2 = img_masks$rimB),
      img_cutoffs
    )
    
    

featureList <- list(
      "manualFeatures" = manualFeatures,
      "M.Blue_Its.Blue" = computeFeatures(x = img_masks$B, ref = img$B), ##Mask: Blue, Ref: Blue, Information of nucleus properties
      "M.Blue_Its.Green" = computeFeatures(x = img_masks$B, ref = img$G), ##Mask: Blue, Ref: Green, Gain: Average signal properties per nucleus
      "M.G.Nuc.Otsu_Its.Green" = computeFeatures(x = img_masks_G.Otsu$Nucleus, ref = img$G), ##Mask: HDR Green in Nucleus, Ref: Green, Gain: Green signal properties in Nucleus
      "M.G.Nuc.Huayu_Its.Green" = computeFeatures(x = img_masks_G.Huayu$Nucleus, ref = img$G), ##See above
      "M.G.Cyt.Otsu_Its.Green" = computeFeatures(x = img_masks_G.Otsu$Cytosol, ref = img$G),##Mask: HDR Green in Cytosol, Ref: Green, Gain: Green signal properties in Cytosol
      "M.G.Cyt.Huayu_Its.Green" = computeFeatures(x = img_masks_G.Huayu$Cytosol, ref = img$G), ##See above
      "M.G.Rest.Otsu_Its.Green" = computeFeatures(x = img_masks_G.Otsu$Rest, ref = img$G),##Mask: Green in Rest, Ref: Green, Gain: Green signal properties in Rest
      "M.G.Cyt.Otsu_Its.Red" = computeFeatures(x = img_masks_G.Otsu$Cytosol, ref = img$R),##Mask: HDR Green in Cytosol, Ref: Red, Gain: Relation of green signal and red signal
      "M.G.Cyt.Huayu_Its.Red" = computeFeatures(x = img_masks_G.Huayu$Cytosol, ref = img$R),##See above
      "M.G.Cyt.Otsu_Its.Yellow" = computeFeatures(x = img_masks_G.Otsu$Cytosol, ref = img$Y),##Mask: HDR Green in Cytosol, Ref: Yellow, Gain: Relation of green signal and yellow signal
      "M.G.Cyt.Huayu_Its.Yellow" = computeFeatures(x = img_masks_G.Huayu$Cytosol, ref = img$Y)##See above
    )  

```

The featureList objects contain all colocalization, shape and texture features. 

```{r}
str(featureList)
```

# SessionInfo

```{r}
sessionInfo()
```

